---
title: "Bay Delta Figure Gallery"
author: "DUWG"
date: "2024-01-12"
output: 
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


This document provides examples and code for commonly-plotted data types using data collected in the San Francisco Estuary.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(viridis)
library(readr)
library(here)
```

# Environmental time series

## Multiple parameters 

```{r RVB-fig, fig.cap = "Air Temperature, Dissolved Oxygen, Electrical Conductivity and Turbidity Data at Rio Vista Bridge (RVB) in 2017. Data were downloaded from CDEC (cdec.water.ca.gov)", fig.width = 8, fig.height = 7.5}
# Read in data
RVB_df <- readRDS(here("data/raw/CDEC_parameters_RVB_2015_2019.rds")) %>%
  filter(!(parameter == "EC" & parameter_value <100))

# Filter to 2017
RVB_2017 <- RVB_df %>% 
  filter(year == 2017) %>%
  mutate(parameter = case_when(parameter == "AirTempF" ~ "Air Temp (째F)",
                               parameter == "DO" ~ "DO (mg/L)",
                               parameter == "EC" ~ "EC (uS/cm)",
                               parameter == "TurbidityNTU" ~ "Turbidity (NTU)"))

# Plot data 
ggplot(data = RVB_2017) + 
  geom_point(aes(x = datetime, y = parameter_value)) + 
  facet_wrap(~parameter, scale = "free_y", nrow = 4, strip.position = "left") +
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b") +
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_blank(),
        strip.background = element_rect(fill="White", color = "White"),
        strip.placement = "outside")
```
## Multiple stations, water year type

```{r CDECstations-fig, fig.cap = "Water Temperature at VER (Vernalis), JER (Jersey Point) and RVB (Rio Vista Bridge) between 2015-2018. Data were downloaded from CDEC (cdec.water.ca.gov). A threshold line of 70째F (approximately 21째C) is displayed on the plots to indicate stressful temperatures for some native fishes.", fig.width = 8, fig.height = 7.5}
# Read in data
watertemp_df <- readRDS(here("data/raw/CDEC_watertemp_JER_VER_RVB_2015_2019.rds")) 
wytype <- read.csv(here("data/raw/WYType.csv")) %>%
  filter(Basin == "SacramentoValley") %>%
  select(WY, Yr.type)

# Filter to 2017
watertemp <- watertemp_df %>%
  rename(station = location_id,
         watertemp = parameter_value) %>%
  mutate(year = year(datetime),
         yearF = factor(year),
         month = month(datetime),
         WY = if_else(month>10, year + 1, year))%>%
  filter(WY <2019 & WY > 2014) %>%
  left_join(wytype, by = "WY")

# Plot data 
ggplot(data = watertemp) + 
  geom_point(aes(x = datetime, y = watertemp, color = Yr.type), size = 1.5) + 
  geom_hline(yintercept = 70, linetype = "dashed") + 
  facet_wrap(~station, nrow = 3) +
  scale_x_datetime(date_breaks = "3 months", date_labels = "%b %Y") +
  labs(y = "Water Temperature (째F)", color = "WY Type") +
  scale_color_viridis(discrete = TRUE)+ 
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

```

# Fish Data

## Fish cpue vs environmental data 

## Abundance index barplot or escapement barplot (year by abundance)

## Salvage by year

## cpue by region (zoop or fish)

## cpue by water year type  

## cpue by action period/inundation period 

* shaded periods of time with overlay of cpue or env data time series 



# Spatial data

## Station map

* mapview
* normal

## Fish detections map
