---
title: "Bay Delta Figure Gallery"
author: "DUWG"
date: "2024-01-12"
output: 
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

This document provides examples and code for commonly-plotted data types using data collected in the San Francisco Estuary.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(viridis)
library(readr)
library(here)
```

# Environmental time series

## Multiple parameters 

```{r RVB-fig, fig.cap = "Air Temperature, Dissolved Oxygen, Electrical Conductivity and Turbidity Data at Rio Vista Bridge (RVB) in 2017. Data were downloaded from CDEC (cdec.water.ca.gov)", fig.width = 8, fig.height = 7.5}
# Read in data
RVB_df <- readRDS(here("data/raw/CDEC_parameters_RVB_2015_2019.rds")) %>%
  filter(!(parameter == "EC" & parameter_value <100))

# Filter to 2017
RVB_2017 <- RVB_df %>% 
  filter(year == 2017) %>%
  mutate(parameter = case_when(parameter == "AirTempF" ~ "Air Temp (째F)",
                               parameter == "DO" ~ "DO (mg/L)",
                               parameter == "EC" ~ "EC (uS/cm)",
                               parameter == "TurbidityNTU" ~ "Turbidity (NTU)"))

# Plot data 
(rvb_plot <- ggplot(data = RVB_2017) + 
  geom_point(aes(x = datetime, y = parameter_value)) + 
  facet_wrap(~parameter, scale = "free_y", nrow = 4, strip.position = "left") +
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b") +
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_blank(),
        strip.background = element_rect(fill="White", color = "White"),
        strip.placement = "outside"))

```

## Export figure

```{r, echo = FALSE, eval = FALSE}
png(here("figures/plot_RVB_2017.png"), units = "in", width = 6, height = 8, res = 300)
rvb_plot
dev.off()
```



## Multiple stations, water year type

```{r CDECstations-fig, fig.cap = "Water Temperature at VER (Vernalis), JER (Jersey Point) and RVB (Rio Vista Bridge) between 2015-2018. Data were downloaded from CDEC (cdec.water.ca.gov). A threshold line of 70째F (approximately 21째C) is displayed on the plots to indicate stressful temperatures for some native fishes.", fig.width = 8, fig.height = 7.5}
# Read in data
watertemp_df <- readRDS(here("data/raw/CDEC_watertemp_JER_VER_RVB_2015_2019.rds")) 
wytype <- read.csv(here("data/raw/WYType.csv")) %>%
  filter(Basin == "SacramentoValley") %>%
  select(WY, Yr.type)

# Filter to 2017
watertemp <- watertemp_df %>%
  rename(station = location_id,
         watertemp = parameter_value) %>%
  mutate(year = year(datetime),
         yearF = factor(year),
         month = month(datetime),
         WY = if_else(month>10, year + 1, year))%>%
  filter(WY <2019 & WY > 2014) %>%
  left_join(wytype, by = "WY")

# Plot data 
ggplot(data = watertemp) + 
  geom_point(aes(x = datetime, y = watertemp, color = Yr.type), size = 1.5) + 
  geom_hline(yintercept = 70, linetype = "dashed") + 
  facet_wrap(~station, nrow = 3) +
  scale_x_datetime(date_breaks = "3 months", date_labels = "%b %Y") +
  labs(y = "Water Temperature (째F)", color = "WY Type") +
  scale_color_viridis(discrete = TRUE)+ 
  theme_bw() + 
  theme(axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

```

# Fish Data

## Abundance index barplot or escapement barplot (year by abundance)

```{r}
FMWTindices <- read_csv("https://filelib.wildlife.ca.gov/Public/TownetFallMidwaterTrawl/FMWT%20Data/FMWTindices.csv") %>%
  janitor::clean_names()

ggplot(FMWTindices) + 
  geom_col(aes(x = factor(year), y = delta_smelt), fill = "steelblue4") +
  geom_text(aes(x = factor(year), y = delta_smelt +100, label = delta_smelt, angle = 90)) + 
  labs(y = "Delta Smelt Index") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) 
```

## DJFMP Data Examples

### Sample size tile plot

```{r}
fishdata <- readRDS(here("data/processed/djfmp_data_2015-2019.rds")) %>%
  mutate(fYear = factor(Year),
         fMonth = factor(Month),
         RegionCode = factor(RegionCode))

unique_samples <- fishdata %>%
  select(-ForkLength, -Count, -IEPFishCode) %>%
  distinct() %>%
  group_by(fYear, fMonth, RegionCode) %>%
  summarize(n = n()) %>%
  ungroup()

ggplot(unique_samples) + 
  geom_tile(aes(fYear, fMonth, fill = n), color = "black") +
  facet_wrap(~RegionCode) +
  labs(title = "Sample Size by Month, Year, Region") +
  scale_fill_viridis(option = "plasma") + 
  theme_classic()
```
### Fish cpue by year and month

```{r}
native_spp <- c("SACSUC", "HITCH", "SPLITT", "SACPIK")
common_spp <- c("MISSIL", "BLUEGI", "WESMOS")

daily_data <- fishdata %>%
  filter(IEPFishCode %in% native_spp) %>%
  mutate(CPUE = Count/Volume) %>%
  group_by(SampleDate, StationCode) %>%
  mutate(dailyCPUE = sum(CPUE)) %>%
  ungroup() %>%
  select(-Datetime, -ForkLength, -Count, -CPUE) %>%
  distinct()

monthly_data <- daily_data %>%
  group_by(fYear, fMonth, IEPFishCode) %>%
  summarize(meanCPUE = mean(dailyCPUE),
            meanTemp = mean(WaterTemp))

ggplot(monthly_data) +
  geom_col(aes(factor(fMonth), meanCPUE, fill = factor(fYear))) +
  facet_grid(factor(fYear)~IEPFishCode, scales = "free_y") +
  scale_fill_viridis(discrete = TRUE) +
  theme_bw()


```

## Salvage by year
```{r}

```


## cpue by region

```{r}
daily_data_all <- fishdata %>%
  mutate(CPUE = Count/Volume) %>%
  select(-Datetime, -ForkLength, -IEPFishCode, -Count) %>%
  distinct() %>%
  group_by(SampleDate, StationCode) %>%
  mutate(dailyCPUE = sum(CPUE, na.rm = TRUE)) %>%
  ungroup()%>%
  select(-CPUE) %>%
  distinct() 

monthly_data_all <- daily_data_all %>%
  group_by(fYear, fMonth, Month, Year, RegionCode) %>%
  summarize(meanCPUE = mean(dailyCPUE)) %>% 
  mutate(WY=if_else(Month>=10, Year + 1,Year))

annual_data_all <- daily_data_all %>%
  group_by(fYear,RegionCode) %>%
  summarize(meanCPUE = mean(dailyCPUE))

ggplot(monthly_data_all) + 
  geom_col(aes(x = RegionCode, y = meanCPUE, fill = fMonth)) + 
  facet_wrap(~fYear) + 
  labs(title = "Mean Monthly CPUE across regions, 2015-2019") +
  scale_fill_viridis(discrete = TRUE)
```

## cpue by water year type  

```{r}
wytype <- read_csv(here::here("data/processed/wytype_1906_cur.csv")) %>%
  rename(WY = Year)

cpue_wytype <- left_join(monthly_data_all, wytype, by = "WY") 
  group_by(WY) 
  

ggplot(cpue_wytype) + 
  geom_col(aes(x = WY, y = meanCPUE, fill = Sac_WYtype)) + 
  scale_fill_viridis(discrete = TRUE) +
  theme_bw() 
```


## cpue by action period/inundation period 

* shaded periods of time with overlay of cpue or env data time series 



# Spatial data

## Station map

* mapview
* normal

```{r}
left_join()
```


## Fish detections map
