---
title: "Rosies Plots"
author: "Rosie"
date: "2024-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(tidyverse)
library(rvest)
library(xml2)

```
# Water Year Index

I'm going to try and graph the water year indices by pulling them from online. Wish me luck.
https://cdec.water.ca.gov/reportapp/javareports?name=WSIHIST 

```{r}


#Read in teh website

wytype <- read_html("https://cdec.water.ca.gov/reportapp/javareports?name=WSIHIST")

#Extract th block of text with the water year type
wytype2 <- wytype %>% 
  html_elements("pre") %>% 
  html_text2() %>% 
  str_split("\r\n") %>% 
  unlist() %>% 
  str_trim() %>% 
  .[-c(1:14)] 

#Choose the rows starting with years, just the first table with the Sac/SJ valley index and Water Year Type  
wytype2_keep <- wytype2[if_else(cumsum(if_else(wytype2 == "", 1, 0)) < 1, TRUE, FALSE)]

#Set up the names for the table
wytype_names <- c(
  "Year", 
  "Sac_OctMar",
  "Sac_AprJul",
  "Sac_WYsum",
  "Sac_Index",
  "Sac_WYtype",
  "SJ_OctMar",
  "SJ_AprJul",
  "SJ_WYsum",
  "SJ_Index",
  "SJ_WYtype"
)

#the first few years only have san joaquin index, so do those seperately
df_wytype_1901_1905 <- read_table(wytype2_keep[1:5], col_names = wytype_names[c(1, 7:11)])

#now parse the rest of the table into coloums
df_wytype_1906_cur <- read_table(wytype2_keep[6:length(wytype2_keep)], col_names = wytype_names)

#Bind the two tables togeter, arrange by water year, and turn the WYtypes into factors with the right orter
df_wytype <- bind_rows(df_wytype_1906_cur, df_wytype_1901_1905) %>% arrange(Year) %>%
  mutate(Sac_WYtype = factor(Sac_WYtype, levels = c("C", "D", "BN", "AN", "W")),
         SJ_WYtype = factor(SJ_WYtype, levels = c("C", "D", "BN", "AN", "W")))

#plot water year index by year, color-coding by water year type. 
ggplot(df_wytype, aes(x = Year, y = Sac_Index, fill = Sac_WYtype)) + geom_col()+
  scale_fill_manual(values = c("firebrick", "orange", "yellow","skyblue", "blue"), name = "Water Year Type")+
  ylab("Sacramento Valley Index")+ theme_bw()

```
Now the san joaquin

```{r}

ggplot(df_wytype, aes(x = Year, y = SJ_Index, fill = SJ_WYtype)) + geom_col()+
  scale_fill_manual(values = c("firebrick", "orange", "yellow","skyblue", "blue"), name = "Water Year Type")+
  ylab("San Joaquin Valley Index")+ theme_bw()


```
# Delta Outflow

There are several different places to get data on net Delta outflow. The first, is CDEC station DTO. This is good for the most recent data, but has issues sometimes.


```{r}
#you can use the 'cder' package to pull the data
library(cder)

DTO = cdec_query("DTO", 23, Duration = "D", start.date = as.Date("2015-01-01"), end.date = as.Date("2023-12-31"))

ggplot(DTO, aes(x = DateTime, y = Value))+ geom_line()+
  ylab("Net Delta Outflow Index (cfs, cdec)")

```

The other data source for Delta Outflow (as well as a lot of other flow parameters) is DWR's Dayflow model. This model is run once per year, usually in January or February of the following water year. It's run after all of the input data has been fully QAQC'd, so you have to wait for it, but the data are better.

It's hosted on the CNRA open data portal and the format is terrible.

https://data.cnra.ca.gov/dataset/dayflow 

```{r}

#download the data

Dayflow2022 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/5300e2e6-8ff0-45b7-ad70-9e0c37ab8fcf/download/dayflowcalculations2022.csv")
  
Dayflow2021 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/83122ce7-e7f5-4ad1-b5e9-6a7032cb1117/download/dayflowcalculations2021.csv")

#the older data uses a different formula for some of the parameters, so they have different names. read the documentation for details on the differences.
Dayflow29_39 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/ab12e85f-82f4-4723-9973-deeed41b2057/download/dayflow-results-1929-1939.csv") %>%
  rename(YOLO = YOLO1, OUT = OUT1, RIO = RIO1, DIVER = DIVER1, EFFDIV = EFFDIV1, WEST = WEST1, Mo = Month) %>%
  mutate(Date = mdy(Date), DIVER = as.numeric(DIVER), EFFDIV = as.numeric(EFFDIV))

Dayflow40_49 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/bf58c67c-63b4-47d4-9a25-2b95e5479a0c/download/dayflow-results-1940-1949.csv")%>%
  rename(YOLO = YOLO2, OUT = OUT2, RIO = RIO2, DIVER = DIVER2, EFFDIV = EFFDIV2, WEST = WEST2, TOT = TOT2)%>%
  mutate(Date = mdy(Date))

Dayflow50_55 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/9225dbe7-54a6-4466-b360-e66f51407683/download/dayflow-results-1950-1955.csv")%>%
  rename(YOLO = YOLO2, OUT = OUT2, RIO = RIO2, DIVER = DIVER2, EFFDIV = EFFDIV2, WEST = WEST2, TOT = TOT2)%>%
  mutate(Date = mdy(Date))

Dayflow56_69 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/3109f3ef-b77b-4288-9ece-3483899d10da/download/dayflow-results-1956-1969.csv")%>%
  mutate(Date = mdy(Date))

Dayflow70_83 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/a0a46a1d-bec5-4db9-b331-655e306860ba/download/dayflow-results-1970-1983.csv")%>%
  mutate(Date = mdy(Date))

Dayflow84_96 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/cb04e626-9729-4105-af81-f6e5a37f116a/download/dayflow-results-1984-1996.csv")%>%
  mutate(Date = mdy(Date))

Dayflow97_20 = read_csv("https://data.cnra.ca.gov/dataset/06ee2016-b138-47d7-9e85-f46fae674536/resource/21c377fe-53b8-4bd6-9e1f-2025221be095/download/dayflow-results-1997-2020.csv")%>%
  mutate(Date = mdy(Date))


Dayflow = bind_rows(Dayflow29_39, Dayflow40_49, Dayflow50_55, Dayflow56_69, Dayflow70_83, Dayflow84_96, Dayflow97_20,
                    Dayflow2021, Dayflow2022)

```

Now let's graph outflow!

```{r}
ggplot(Dayflow, aes(x = Date, y = OUT)) + geom_line()
```
X2 is also included, at least 1995-2022. If you want older X2 estimates, check out - 

Hutton, P. H., J. S. Rath, L. Chen, M. J. Ungs, and S. J. Roy. 2016. Nine Decades of Salinity Observations in the San Francisco Bay and Delta: Modeling and Trend Evaluations. J. Water Resour. Plann. Mange. [accessed 2024 Jan 23].  142 (3):04015069. https://doi.org/10.1061/(ASCE)WR.1943-5452.0000617. 


```{r}
ggplot(filter(Dayflow, !is.na(X2)), aes(x = Date, y = X2))+ geom_line()
```

